A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-04 paper.Rmd
working on section 5.6

The prediction in a regression model is $E(Y|X)$, that is, it is conditional on observed values of the predictors. The distribution of $X$, or consequently $\hat{Y}$, may however affect the ability to read any patterns in the residuals. The four distributions of fitted values, used in this experiment, were designed to examine this.

Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of fitted value distribution on the power of conventional tests and visual tests. For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown. For visual tests, note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will affect the power of the visual test. To have a fair comparison, for lineups with uniform fitted value distribution, we randomly sample five evaluations from the total eleven evaluations to estimate the power curves. The sampling process is repeated 500 times to produce an indication of the variation of the power.

@@

```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests. Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations. The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines. The fitted value distribution has greater impact on the power of conventional tests than visual tests. Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450}

@@

>>>>

In regression predictions are conditional on the observed values of the predictors, that is, the conditional mean of the dependent variable Y given the value of the independent variable X, $E(Y|X)$. This is an often forgotten element of regression analysis but it is important. Where X is observed, particularly the distribution of the X values in the sample, or consequently $\hat{Y}$, may affect the ability to read any patterns in the residual plots. This experiment was constructed to assess this, based on four different distributions of fitted values: uniform, normal, discrete and lognormal (skewed). We would expect that if $\hat{Y}$ has a uniform distribution, this would make it easier to read the relationship with the residuals.

@@

```{r different-x-dist-poly-power, fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests. Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations. The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines. The fitted value distribution has greater impact on the power of conventional tests than visual tests. Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.", dev = 'png', dpi = 450, fig.pos="t"}

@@

Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional and visual tests. For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown. For visual tests, note that more evaluations on lineup with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the total eleven evaluations to estimate the power curves. Re-sampling produces the multiple curves for the uniform, and provides an indication of the variation of the power.

### 2023-04-04 paper.tex
working on section 5.6

The prediction in a regression model is \(E(Y|X)\), that is, it is
conditional on observed values of the predictors. The distribution of
\(X\), or consequently \(\hat{Y}\), may however affect the ability to
read any patterns in the residuals. The four distributions of fitted
values, used in this experiment, were designed to examine this.

Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of
fitted value distribution on the power of conventional tests and visual
tests. For conventional tests, only the power curves of appropriate
tests which are RESET tests and Breusch-Pagan tests for non-linearity
pattern and heteroskedasticity pattern respectively are shown. For
visual tests, note that we have collected more evaluations on lineup
with uniform fitted value distribution, and the number of evaluations on
a lineup will affect the power of the visual test. To have a fair
comparison, for lineups with uniform fitted value distribution, we
randomly sample five evaluations from the total eleven evaluations to
estimate the power curves. The sampling process is repeated 500 times to
produce an indication of the variation of the power.

>>>>

In regression predictions are conditional on the observed values of the
predictors, that is, the conditional mean of the dependent variable Y
given the value of the independent variable X, \(E(Y|X)\). This is an
often forgotten element of regression analysis but it is important.
Where X is observed, particularly the distribution of the X values in
the sample, or consequently \(\hat{Y}\), may affect the ability to read
any patterns in the residual plots. This experiment was constructed to
assess this, based on four different distributions of fitted values:
uniform, normal, discrete and lognormal (skewed). We would expect that
if \(\hat{Y}\) has a uniform distribution, this would make it easier to
read the relationship with the residuals.

\begin{figure}[t]
